---
  output: github_document
---
  
<!-- README.md is generated from README.Rmd. Please edit that file -->
  
```{r, include = FALSE}
require(knitr)
knitr::opts_knit$set(global.par = TRUE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "500px",
  dpi = 150,
  fig.align = "center",
  fig.width=5,
  fig.height=5
)
```

```{r echo=FALSE}
par(mar=c(5,5,1,1))
```
# ggfree: ggplot2-style plots with just base R graphics

[ggplot2](https://ggplot2.tidyverse.org/) is a popular R graphics package that is becoming synonymous with data visualization in R.
The community of developers working within the ggplot2 framework have implemented some [rather nice extensions](https://www.ggplot2-exts.org/gallery/) as well.
However, I personally find the ggplot2 syntax to be confusing, most probably because I've been using the base graphics package in R to generate plots for over a decade.
I seldom see a ggplot2-based visualization that couldn't also be implemented in base graphics.

The overall purpose of `ggfree` is to make it easier to generate plots in the style of [ggplot2](https://ggplot2.tidyverse.org/) and its extensions, without ever actually using any ggplot2 code.


## Installation

For the time being, the simplest method is to use the `devtools` package:
```R
# install.packages('devtools')  # if you haven't already installed devtools
require(devtools)
devtools::install_github("ArtPoon/ggfree")
```

(You could also simply copy-paste the functions you want to use into your R script.)



## Examples

### Fancy gridlines

The major and minor gridlines on a shaded background is one of the distinctive features of the default ggplot2 theme.
This implementation uses `rect` and `abline` to obtain the same look.
```{r}
require(ggfree)
# prepare the plot region
plot(NA, xlim=range(faithful$eruptions), ylim=range(faithful$waiting),
     xlab='Duration of eruption (mins)', ylab='Waiting time (mins)')
add.grid()  # <-- a ggfree function!
points(faithful$eruptions, faithful$waiting, pch=21, bg='white')
```

We can also vary the colours, turn off major or minor gridlines, or skip vertical or horizontal lines:
```{r}
plot(NA, xlim=range(faithful$eruptions), ylim=range(faithful$waiting),
     xlab='Duration of eruption (mins)', ylab='Waiting time (mins)')
# now with non-default arguments!
add.grid(mode='x', bg.col='seashell2', lwd.minor=0)
points(faithful$eruptions, faithful$waiting, pch=21, bg='white')
```


### Slopegraphs

In general, a slopegraph is a method for visually comparing a paired set of observations.
To illustrate the use of slopegraphs, I've adapted a carbon dioxide emissions data set similar to the one used by @clauswilke to demonstrate slopegraphs in *Fundamentals of Data Visualization*.
This data sets is packaged with `ggfree`:
```{r}
require(ggfree)
co2.emissions
```

There are two styles of slopegraphs that are implemented in `ggfree`.
For the first style (where the argument `type` is left at its default value `'b'`), a vertical axis provides reference for the numerical values:
```{r}
slopegraph(co2.emissions, colorize=T)
```
Setting `colorize` to `TRUE` causes the line segments to be coloured to emphasize positive and negative slopes.

For the second style (setting `type='t'`), we substitute the raw numerical values for data points, which makes the vertical axis unnecessary:
```{r}
par(family='Palatino')  # use a fancier font

# the actual code here
slopegraph(co2.emissions, type='t', cex.text=0.6, names.arg=c(2000, 2010))

# make a nice title
title(expression(text=paste('CO'[2], ' emissions (metric tons) per capita')), 
      cex=0.8)
```


### Ringplots

A ring- or donut-plot is simply a piechart with a hole in it.
Piecharts have been criticized for being potentially misleading because the quantity represented by a given sector ("pie slice") is proportional to its arc (pie crust) length, and not the total area, and because it is difficult to compare two sectors by eye.


### Ridgeplots




